---
title: "p8105_hw2_tj2383"
author: "Tanvi Jain"
date: "9/25/2018"
output: github_document
---

# Homework Two

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

Read in and clean the transit dataset:
```{r}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, north_south_street, east_west_street, ada) %>% mutate(entry_logical = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

This dataset was first imported using the read_csv statement. I then cleaned the variable names so that they were converted to lower case snake, this makes the variables easy to use while coding. I further cleaned the dataset by selecting key variables to include in the dataset; line, station name, station latitude, station longitude, routes, entry, vending, north-south entrance, east-west entrance, and ADA compliance. Lastly, I changed the "entry" variable from a character variable to a logical variable and called that new variable "entry_logical". The dimensions of this dataset are `r dim(transit_data)`. This dataset is not tidy. For example, there are several "route" variables that could be consolidated.





