---
title: "p8105_hw2_tj2383"
author: "Tanvi Jain"
date: "9/25/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

Read in and clean the transit dataset:
```{r cleaning}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", col_types = "cccddcccccccccccccccccccccccddcc") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

According to the cleaning code chunk, this dataset was first imported using the read_csv statement. I then cleaned the variable names so that they were converted to lower case and the snake format, this makes the variables easy to use while coding. I further cleaned the dataset by selecting key variables to include in the dataset; line, station name, station latitude, station longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, and ada. Lastly, I changed the "entry" variable from a character variable to a logical variable. The dimensions of this dataset are `r dim(transit_data)`. This dataset is still not tidy. For example, there are several "route" variables that could be consolidated.


The following were conducted in this code chunk:
-How many distinct stations are there?
-How many stations are ADA compliant?
-What proportion of station entrances / exits without vending allow entrance?
-Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? How many are ADA compliant?
```{r problem1_q}
transit_data %>% 
  distinct(line, station_name) 
select(transit_data, line, station_name, ada) %>% 
  filter(ada == "TRUE") %>% 
  distinct(line, station_name, ada)
select(transit_data, entry, vending) %>% 
  filter(vending == "NO") %>% 
  summarize(mean(entry))
```

According to the problem1_q code chunk, there are `r count(distinct(transit_data, line, station_name))` distinct stations. There are `r count(select(transit_data, line, station_name, ada) %>% 
  filter(ada == "TRUE") %>% 
  distinct(line, station_name, ada))` ADA compliant stations. The proportion of station entrace/exits with no vending are `r select(transit_data, entry, vending) %>% 
  filter(vending == "NO") %>% 
  summarize(mean(entry))`.




